\TheTitle

### 2019-07-07 Paper1/adaptivity.tex
typos corrected

Now we need to specify how the different building blocks of the IPM method can be modified to work with varying truncation orders on different cells. Let us first add dimensions to the notation of the dual iteration function $\bm d$, which has been defined in \eqref{eq:dualIterationFunction}. Now, we have 
$\bm{d}_{\ell}:\mathbb{R}^{N_{\ell}\times p}\times\mathbb{R}^{N_{\ell}\times p}\to\mathbb{R}^{N_{\ell}\times p}$, given by
\begin{align}\label{eq:dualIterationFunctionAd}
\bm{d}_{\ell}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}_{\ell}(\bm{\lambda})\cdot \left(\langle \bm u_{s}(\bm{\lambda}^T\bm{\varphi}_{\ell})\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T-\bm{\hat{u}}\right),
\end{align}
where $\bm{\varphi}_{\ell}\in\mathbb{R}^{N_{\ell}}$ collects all basis functions with total degree smaller or equal to $\ell$. The preconditioner $\bm{B}_{\ell}$ is given by 
\begin{align*}
\bm{B}_{\ell}(\bm{\lambda}) := \langle \nabla \bm{u}_{\bm{s}} (\bm{\lambda}^T\bm{\varphi}_{\ell})\bm{\varphi}_{\ell}\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^{-T}.
\end{align*}
An adaptive version of the moment iteration \eqref{eq:momentIterationFunction} is denoted by $\bm c_{\ell}^{\bm{\ell}'}:\mathbb{R}^{N_{\ell_1'}\times p}\times \mathbb{R}^{N_{\ell_2'}\times p}\times \mathbb{R}^{N_{\ell_3'}\times p}\rightarrow \mathbb{R}^{N_{\ell}\times p}$ and given by
\begin{align*}
\bm{c}_{\ell}^{\bm{\ell}'}\left(\bm{\lambda}_{1},\bm{\lambda}_2,\bm{\lambda}_3\right):= \langle \bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'})\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T - \frac{\Delta t}{\Delta x}\left(\langle \bm g(\bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'}),\bm u_{s}(\bm{\lambda}_3^T\bm{\varphi}_{\ell_3'}))\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T-\langle \bm g(\bm u_{s}(\bm{\lambda}_{1}^T\bm{\varphi}_{\ell_1'}),\bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'}))\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T\right).
\end{align*}
Hence, the index vector $\bm\ell'\in\mathbb{N}^{3}$ denotes the refinement levels of the stencil cells, which are used to compute the time updated moment vector at level $\ell$.

The strategy now is to perform the dual update for a set of moment vectors $\bm{\hat u}_j^n$ at refinement levels $\ell_j^n$ for $j = 1,\cdots,N_x$. Hence, the dual iteration makes use of the iteration function \eqref{eq:dualIterationFunctionAd} at refinement level $\ell_j^n$. After that, the refinement level at the next time step $\ell_j^{n+1}$ is determined by making use of the smoothness indicator \eqref{eq:errorIndicator}. The moment update then computes the moments at the time updated refinement level $\ell_j^{n+1}$ making use of the dual states at the old refinement levels $\bm{\ell}' = (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$. The IPM algorithm with adaptivity then reads
\begin{algorithm}[H]
\begin{algorithmic}[1]
\For{$j=0$ to $N_x+1$}
\State $\ell_j^0 \leftarrow$ choose initial refinement level
\State $\bm{u}_j^0 \leftarrow \frac{1}{\Delta x} \int_{x_{j-1/ 2}}^{x_{j+1/ 2}} \langle u_{\text{IC}}(x, \cdot) \bm{\varphi}_{\ell_j^0} \rangle_{Q_{\ell_j^0}} dx$
\EndFor
\For{$n=0$ to $N_t$}
\For{$j=0$ to $N_x+1$}
\State $\bm{\lambda}_j^{(0)} \leftarrow \bm{\hat \lambda}_j^{n}$
\While{\eqref{eq:tauCrit} is violated}
\State $\bm{\lambda}_j^{(m+1)} \leftarrow \bm{d}_{\ell_j^n}(\bm{\lambda}_{j}^{(m)};\bm{\hat u}_j^{n})$
\State $m \leftarrow m+1$
\EndWhile
\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(m)}$
\State $\ell_j^{n+1}\leftarrow \text{DetermineRefinementLevel}\left(\bm{\hat \lambda}_j^{n+1}\right)$
\EndFor
\For{$j=1$ to $N_x$}
\State $\bm\ell' \leftarrow (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$
\State $\bm{\hat u}_j^{n+1} \leftarrow \bm{c}_{\ell_j^{n+1}}^{\bm\ell'}(\bm{\hat \lambda}_{j-1}^{n+1},\bm{\hat \lambda}_j^{n+1},\bm{\hat \lambda}_{j+1}^{n+1})$
\EndFor
\EndFor
\end{algorithmic}
\caption{Adaptive IPM implementation}
\label{alg:ad-IPM}
\end{algorithm}

>>>>

Now we need to specify how the different building blocks of the IPM method can be modified to work with varying truncation orders in different cells. Let us first add dimensions to the notation of the dual iteration function $\bm d$, which has been defined in \eqref{eq:dualIterationFunction}. Now, we have 
$\bm{d}_{\ell}:\mathbb{R}^{N_{\ell}\times p}\times\mathbb{R}^{N_{\ell}\times p}\to\mathbb{R}^{N_{\ell}\times p}$, given by
\begin{align}\label{eq:dualIterationFunctionAd}
\bm{d}_{\ell}(\bm{\lambda},\bm{\hat{u}}):= \bm{\lambda}-\bm{B}_{\ell}(\bm{\lambda})\cdot \left(\langle \bm u_{s}(\bm{\lambda}^T\bm{\varphi}_{\ell})\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T-\bm{\hat{u}}\right),
\end{align}
where $\bm{\varphi}_{\ell}\in\mathbb{R}^{N_{\ell}}$ collects all basis functions with total degree smaller or equal to $M_\ell$. The preconditioner $\bm{B}_{\ell}$ is given by 
\begin{align*}
\bm{B}_{\ell}(\bm{\lambda}) := \langle \nabla \bm{u}_{\bm{s}} (\bm{\lambda}^T\bm{\varphi}_{\ell})\bm{\varphi}_{\ell}\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^{-T}.
\end{align*}
An adaptive version of the moment iteration \eqref{eq:momentIterationFunction} is denoted by $\bm c_{\ell}^{\bm{\ell}'}:\mathbb{R}^{N_{\ell_1'}\times p}\times \mathbb{R}^{N_{\ell_2'}\times p}\times \mathbb{R}^{N_{\ell_3'}\times p}\rightarrow \mathbb{R}^{N_{\ell}\times p}$ and given by
\begin{align*}
\bm{c}_{\ell}^{\bm{\ell}'}\left(\bm{\lambda}_{1},\bm{\lambda}_2,\bm{\lambda}_3\right):= \langle \bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'})\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T - \frac{\Delta t}{\Delta x}\left(\langle \bm g(\bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'}),\bm u_{s}(\bm{\lambda}_3^T\bm{\varphi}_{\ell_3'}))\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T-\langle \bm g(\bm u_{s}(\bm{\lambda}_{1}^T\bm{\varphi}_{\ell_1'}),\bm u_{s}(\bm{\lambda}_2^T\bm{\varphi}_{\ell_2'}))\bm{\varphi}_{\ell}^T\rangle_{Q_{\ell}}^T\right).
\end{align*}
Hence, the index vector $\bm\ell'\in\mathbb{N}^{3}$ denotes the refinement levels of the stencil cells, which are used to compute the time updated moment vector at level $\ell$.

The strategy now is to perform the dual update for a set of moment vectors $\bm{\hat u}_j^n$ at refinement levels $\ell_j^n$ for $j = 1,\cdots,N_x$. Hence, the dual iteration makes use of the iteration function \eqref{eq:dualIterationFunctionAd} at refinement level $\ell_j^n$. After that, the refinement level at the next time step $\ell_j^{n+1}$ is determined by making use of the smoothness indicator \eqref{eq:errorIndicator}. The moment update then computes the moments at the time updated refinement level $\ell_j^{n+1}$ making use of the dual states at the old refinement levels $\bm{\ell}' = (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$. The IPM algorithm with adaptivity then reads
\begin{algorithm}[H]
\begin{algorithmic}[1]
\For{$j=0$ to $N_x+1$}
\State $\ell_j^0 \leftarrow$ choose initial refinement level
\State $\bm{u}_j^0 \leftarrow \frac{1}{\Delta x} \int_{x_{j-1/ 2}}^{x_{j+1/ 2}} \langle u_{\text{IC}}(x, \cdot) \bm{\varphi}_{\ell_j^0} \rangle_{Q_{\ell_j^0}} dx$
\EndFor
\For{$n=0$ to $N_t$}
\For{$j=0$ to $N_x+1$}
\State $\bm{\lambda}_j^{(0)} \leftarrow \bm{\hat \lambda}_j^{n}$
\While{\eqref{eq:tauCrit} is violated}
\State $\bm{\lambda}_j^{(m+1)} \leftarrow \bm{d}_{\ell_j^n}(\bm{\lambda}_{j}^{(m)};\bm{\hat u}_j^{n})$
\State $m \leftarrow m+1$
\EndWhile
\State $\bm{\hat \lambda}_j^{n+1} \leftarrow \bm{\lambda}_j^{(m)}$
\State $\ell_j^{n+1}\leftarrow \text{DetermineRefinementLevel}\left(\bm{\hat \lambda}_j^{n+1}\right)$
\EndFor
\For{$j=1$ to $N_x$}
\State $\bm\ell' \leftarrow (\ell_{j-1}^n,\ell_{j}^n,\ell_{j+1}^n)^T$
\State $\bm{\hat u}_j^{n+1} \leftarrow \bm{c}_{\ell_j^{n+1}}^{\bm\ell'}(\bm{\hat \lambda}_{j-1}^{n+1},\bm{\hat \lambda}_j^{n+1},\bm{\hat \lambda}_{j+1}^{n+1})$
\EndFor
\EndFor
\end{algorithmic}
\caption{Adaptive IPM implementation}
\label{alg:ad-IPM}
\end{algorithm}
%TODO Auch adaptiver Algorithmus fuer osIPM?

### 2019-07-07 Paper1/introduction.tex
typos corrected

\bm{u}(t,\bm x,\bm \xi) \approx \mathcal{U}(\bm{\hat u},\cdots,\bm{\hat u}_N;\bm\xi).

@@

\mathcal{U}_{\text{SG}}(\bm{\hat u}_0,\cdots,\bm{\hat u}_N;\bm\xi):= \sum_{|i|\leq M} \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi}(\bm\xi),

@@

@@

In this paper, we present a semi-intrusive code framework, which facilitates the task of implementing general intrusive methods. The framework only requires the numerical flux of the deterministic problem (as well as an entropy if IPM is used). We thereby provide the ability to recycle existing implementations of deterministic solvers.
Furthermore, we investigate intrusive methods for steady problems and compare them to collocation methods. The steady setting provides different opportunities to take advantage of features of intrusive methods: 
\begin{itemize}
\item Accelerate convergence to the IPM steady state solution by applying IPM as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method such as IPM, which we then again converge to steady state. 
\item Compute inexact dual variables \eqref{eq:dualProblem} for IPM: Since the moments during the iteration process are inaccurate, i.e. they are not the correct steady state solution, we propose to not fully converge the dual iteration, which computes \eqref{eq:dualProblem}. 
\end{itemize}
In addition to these two strategies for steady problems, we make us of adaptivity: The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not). We test the effectiveness of these acceleration ideas by comparing results with stochastic-Collocation for the uncertain NACA test case.

The paper is structured as follows: After the introduction, we discuss the numerical discretization as well as the implementation and structure of the semi-intrusive framework in section \ref{sec:framework}. Section \ref{sec:collIPM} discusses the IPM acceleration with a non-intrusive method and in section \ref{sec:OneShotIPM}, we discuss the idea of not converging the dual iteration. A comparison of results computed with the presented methods is given in \ref{sec:results}, followed by a summary and outlook in section \ref{sec:summary_outlook}.

>>>>

\bm{u}(t,\bm x,\bm \xi) \approx \mathcal{U}(\bm{\hat u};\bm\xi).

@@

\mathcal{U}_{\text{SG}}(\bm{\hat u};\bm\xi):= \sum_{|i|\leq M} \bm{\hat{u}}_i\varphi_i(\bm{\xi}) = \hat{\bm u}^T\bm{\varphi}(\bm\xi),

@@

\bm{\hat u}_i(t=0,\bm{x}&) = \langle\bm{u}_{\text{IC}}(\bm{x},\cdot)\varphi_i\rangle.

@@

In this paper, we present a semi-intrusive code framework, which facilitates the task of implementing general intrusive methods. The framework only requires the numerical flux of the deterministic problem (as well as an entropy if IPM is used). We thereby provide the ability to recycle existing implementations of deterministic solvers.
Furthermore, we investigate intrusive methods for steady problems and compare them against collocation methods. The steady setting provides different opportunities to take advantage of features of intrusive methods: 
\begin{itemize}
\item Accelerate convergence to the IPM steady state solution by applying IPM as a post-processing step for collocation methods: We converge the moments of the solution to a steady state with an inaccurate, but cheap collocation method and then use the resulting collocation moments as starting values for an expensive but accurate intrusive method such as IPM, which we then again converge to steady state. 
\item Compute inexact dual variables \eqref{eq:dualProblem} for IPM: Since the moments during the iteration process are inaccurate, i.e. they are not the correct steady state solution, we propose to not fully converge the dual iteration, which computes \eqref{eq:dualProblem}. 
\end{itemize}
In addition to these two strategies for steady problems, we make use of adaptivity: The intrusive nature of SG and IPM can be used to locally increase the number of moments whenever the solution has a complex structure in the random variable (as well as decrease the number of moments if not). We test the effectiveness of these acceleration ideas by comparing results with stochastic-Collocation for the uncertain NACA test case.

The paper is structured as follows: After the introduction, we discuss the numerical discretization as well as the implementation and structure of the semi-intrusive framework in section \ref{sec:framework}. Section \ref{sec:collIPM} discusses the IPM acceleration with a non-intrusive method and in section \ref{sec:OneShotIPM}, we discuss the idea of not converging the dual iteration. Section~\ref{sec:adaptivity} extends the presented numerical framework to an algorithm making use of adaptivity. A comparison of results computed with the presented methods is then given in \ref{sec:results}, followed by a summary and outlook in section \ref{sec:summary_outlook}.

### 2019-07-07 Paper1/one_shot_ipm.tex
paper continued
- N notation further corrections
- quadrature added
- adaptivity continued
typos corrected

In this section we aim at breaking up the inner loop in the IPM algorithm \ref{alg:IPM}, i.e. to just perform one iteration of the dual problem in each time step. Consequently, the IPM reconstruction given by \eqref{eq:primalProblem} is not done exactly, meaning that the reconstructed solution does not minimize the entropy while not fulfilling the moment constraint. However, the fact that the moment vectors are not yet converged to the steady solution seems to permit such an inexact reconstruction. Hence, we aim at iterating the moments to steady state and the dual variables to the exact solution of the IPM optimization problem \eqref{eq:primalProblem} simultaneously.
Hence, by successively performing one update of the moment iteration and one update of the dual iteration, we obtain 
\begin{subequations}\label{eq:oneshotIPM}
\begin{align}
&\bm{\lambda}_{j}^{n+1} =  \bm{d}(\bm{\lambda}_j^{n},\bm{u}_j^{n}) \enskip \text{ for all j} \label{eq:oneshotIPMdual}\\
&\bm{u}_j^{n+1} =  \bm{c}\left(\bm{\lambda}_{j-1}^{n+1},\bm{\lambda}_{j}^{n+1},\bm{\lambda}_{j+1}^{n+1}\right)\label{eq:oneshotIPMmoment}.
\end{align}
\end{subequations}
This gives algorithm
\begin{algorithm}[H]
\begin{algorithmic}[1]
\For{$j=0$ to $N_x+1$}
\State $\bm{u}_j^0 \leftarrow \frac{1}{\Delta x} \int_{x_{j-1/ 2}}^{x_{j+1/ 2}} \langle u_{\text{IC}}(x, \cdot) \bm{\varphi} \rangle dx$
\EndFor
\While{\eqref{eq:residualUnsteady} is violated}
\For{$j=1$ to $N_x$}
\State $\bm{\lambda}_j^{n+1} \leftarrow \bm{d}(\bm{\lambda}_{j}^{n};\bm{\hat u}_j^{n})$
\State $\bm{\hat u}_j^{n+1} \leftarrow \bm{c}(\bm{\lambda}_{j-1}^{n+1},\bm{\lambda}_j^{n+1},\bm{\lambda}_{j+1}^{n+1})$
\EndFor
\State $n \leftarrow n+1$
\EndWhile
\end{algorithmic}
\caption{One-shot IPM implementation}
\label{alg:osIPM}
\end{algorithm}
We call this method One-Shot IPM, since it is inspired by One-shot optimization, see for example \cite{hazra2005aerodynamic}, which uses only a single iteration of the primal and dual step in order to update the design variables. Note that the dual variables from the One-Shot iteration are written without a hat to indicate that they are not the exact solution of the dual problem.

@@

First, understand what contraction of the classical IPM iteration implies, we rewrite the moment iteration \eqref{eq:momentIteration} of the classical IPM scheme: When defining the update function

>>>>

In this section we aim at breaking up the inner loop in the IPM algorithm \ref{alg:IPM}, i.e. to just perform one iteration of the dual problem in each time step. Consequently, the IPM reconstruction given by \eqref{eq:primalProblem} is not done exactly, meaning that the reconstructed solution does not minimize the entropy while not fulfilling the moment constraint. However, the fact that the moment vectors are not yet converged to the steady solution seems to permit such an inexact reconstruction. Hence, we aim at iterating the moments to steady state and the dual variables to the exact solution of the IPM optimization problem \eqref{eq:primalProblem} simultaneously.
Hence, by successively performing one update of the moment iteration and one update of the dual iteration, we obtain 
\begin{subequations}\label{eq:oneshotIPM}
\begin{align}
&\bm{\lambda}_{j}^{n+1} =  \bm{d}(\bm{\lambda}_j^{n},\bm{u}_j^{n}) \enskip \text{ for all j} \label{eq:oneshotIPMdual}\\
&\bm{u}_j^{n+1} =  \bm{c}\left(\bm{\lambda}_{j-1}^{n+1},\bm{\lambda}_{j}^{n+1},\bm{\lambda}_{j+1}^{n+1}\right)\label{eq:oneshotIPMmoment}.
\end{align}
\end{subequations}
This gives algorithm
\begin{algorithm}[H]
\begin{algorithmic}[1]
\For{$j=0$ to $N_x+1$}
\State $\bm{u}_j^0 \leftarrow \frac{1}{\Delta x} \int_{x_{j-1/ 2}}^{x_{j+1/ 2}} \langle u_{\text{IC}}(x, \cdot) \bm{\varphi} \rangle_Q dx$
\EndFor
\While{\eqref{eq:residualUnsteady} is violated}
\For{$j=1$ to $N_x$}
\State $\bm{\lambda}_j^{n+1} \leftarrow \bm{d}(\bm{\lambda}_{j}^{n};\bm{\hat u}_j^{n})$
\State $\bm{\hat u}_j^{n+1} \leftarrow \bm{c}(\bm{\lambda}_{j-1}^{n+1},\bm{\lambda}_j^{n+1},\bm{\lambda}_{j+1}^{n+1})$
\EndFor
\State $n \leftarrow n+1$
\EndWhile
\end{algorithmic}
\caption{One-shot IPM implementation}
\label{alg:osIPM}
\end{algorithm}
We call this method One-Shot IPM, since it is inspired by One-shot optimization, see for example \cite{hazra2005aerodynamic}, which uses only a single iteration of the primal and dual step in order to update the design variables. Note that the dual variables from the One-Shot iteration are written without a hat to indicate that they are not the exact solution of the dual problem.

@@

First, to understand what contraction of the classical IPM iteration implies, we rewrite the moment iteration \eqref{eq:momentIteration} of the classical IPM scheme: When defining the update function
